---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The {ppl.tits} package

<!-- badges: start -->
<!-- badges: end -->

The goal of {ppl.tits} package is to encapsulate all the `R` code used to prepare and analyse the data for the *Tits reproduction study* of the PubPrivLands project. The data generated by the functions of this package are progressively made accessible as the package is developed. To access data or functions, you need first to install this package.  
\


## How to install this package?

You can install the development version of {ppl.tits} with the `{devtools}` package:

``` {r package installation}
install.packages("devtools") # To install the {devtools} package.
devtools::install_github("mrelnoob/ppl.tits") # To install the {ppl.tits} package.
```



## Where can I find the data?

If you want to access the **raw** datasets used in this package, you have to download them from the *mydata/* folder on the [{ppl.tits} project's page](https://github.com/mrelnoob/ppl.tits) of my Github account. You can also contact me directly through: francois-marie.martin@univ-fcomte.fr   

To load the latest version of the **cleaned dataset** directly in your `R` session, you need to call the `tdata_upD_final()` function as in:

```{r data access example}
tits_data <- ppl.tits::tdata_upD_final()$final_dataset
```

For access to other datasets (e.g. raw data, working datasets), please refer to the functions documentation pages:

```{r doc access}
help(package=ppl.tits, try.all.packages = TRUE) # To go to the package's index page allowing
# access to all functions' documentation pages (help pages). Alternatively, you can go to a 
# function's help page directly by using the command `?myfunction` as in:
?tdata_upD_final
```



## Where can I find *progress reports*?

You can find **HTML reports** detailing most data processing and analyses performed in this study in the *output/texts/* folder of the {ppl.tits} project. To access them, you need to download a copy of the project's folder by clicking the green **Code** button (and then **Download ZIP**) on the [{ppl.tits} project's page](https://github.com/mrelnoob/ppl.tits).



## Where can I find the Random Forest object from the *local habitat quality* model?

This `randomForest` object is generated and exported by the `local_quality_model()` function and can be accessed as follows:

```{r rf export}
ppl.tits::local_quality_model()$rf4pm
```



******
TO BE CONTINUED...(ongoing development)
