---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The {ppl.tits} package

<!-- badges: start -->
<!-- badges: end -->

The goal of `{ppl.tits}` package is to encapsulate all the `R` code used to prepare and analyse the data for the *Tits reproduction study* of the PubPrivLands project. The data generated by the functions of this package are progressively made accessible as the package is developed. To access data or functions, you need first to install this package.  
\


## How to install this package?

You can install the development version of `{ppl.tits}` with the `{devtools}` package:

``` {r package installation}
install.packages("devtools") # To install the {devtools} package.
devtools::install_github("mrelnoob/ppl.tits") # To install the {ppl.tits} package.
```



## Where can I find the data?

The **raw tits observations** are an exported dataset (named `raw_tits_data`) of this package and can be directly loaded in an `R` session with the following call:

```{r rawdata access example}
library(ppl.tits)
tits_data <- raw_tits_data
# Or, without loading the package:
tits_data <- ppl.tits::raw_tits_data

# See also:
?raw_tits_data # Currently unavailable (***).
```

If you want to access the other **raw** datasets used in this package (e.g. temperature records, environmental variable computations), you have to download them from the *input_raw_data/* folder on the [{ppl.tits} project Github page](https://github.com/mrelnoob/ppl.tits) by clicking the green **Code** button (and then **Download ZIP**) on the top-right corner of the page.  
You can also contact me directly through: francois-marie.martin@univ-fcomte.fr   

The latest version of the **cleaned nestling dataset** (named `ntits_clean`) is an exported dataset of this package and can be directly loaded in an `R` session with the following call:

```{r data access example}
library(ppl.tits)
tits_data <- ntits_clean
# Or, without loading the package:
tits_data <- ppl.tits::ntits_clean

# For a description of these data, see:
?ntits_clean # Currently incomplete (***documentation in progress).
```

For access to other datasets (e.g. raw data, working datasets), please refer to the functions documentation pages:

```{r doc access}
help(package=ppl.tits, try.all.packages = TRUE) # To go to the package's index page allowing
# access to all functions' documentation pages (help pages). Alternatively, you can go to a 
# function's help page directly by using the command `?myfunction` as in:
?tdata_upD_final
```



## Where can I find *progress reports*?

You can find **HTML reports** detailing most data processing and analyses performed in this study in the *output/texts/* folder of the {ppl.tits} project. To access them, you need to download a copy of the project's folder by clicking the green **Code** button (and then **Download ZIP**) on the [{ppl.tits} Github page](https://github.com/mrelnoob/ppl.tits).



## Where can I find the Random Forest object from the *local habitat quality* model?

This `randomForest` object is generated and exported by the `local_quality_model()` function and can be accessed as follows:

```{r rf export}
rf <- ppl.tits::local_quality_model()$rf4pm
```
\
\

******

# Package building

The {ppl.tits} has been built through a pipeline programming procedure using the {targets} package.
TO BE CONTINUED...
