---
title: "Intermediate analyses report - Tits reproduction study, PubPrivLands project (Martin *et al.*, *in prep.*)"
author: "François-Marie Martin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: ppl.tits_biblio.bib
csl: fanf_style.csl
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# IMPORTANT NOTE: if you read this, then know that this RMarkdown report has been written as a 'target' in a {targets} pipeline (https://books.ropensci.org/targets/). If you simply want to knit this report, you need to do it using using 'targets::tar_make()' (provided that you did not altered the structure of the "ppl.tits" folder to which this report belongs), otherwise it won't work! Alternatively, you can paste this .Rmd at the root of the "ppl.tits" folder. If you want to run the code contained within the code chunks of this document WITHOUT knitting, then you need to set the working directory of this .Rmd document at the root of the "ppl.tits" folder using the global options of RStudio, 'setwd()', or 'knitr::opts_knit$set(root.dir = '~/')'. That is because R, RStudio or {knitr} do not seem to be able to find the "_targets" folder containing the targets that are called by my code chunks otherwise... By default, RMarkdown sets the working directory where the .Rmd file is stored, not at the root of the R project the report is a part of.
# Still, note that theoretically, you do not need to knit or modify this document as it is meant for reproducibility and you can simply read the .html file for that.
```

# 1. Introduction
## 1.1. Regarding reproductibility


In order to facilitate any attempt at reproducing this study, here is a depiction of the system information used to prepare and analyse the data from the PubPrivLands tits reproduction study (Martin *et al.* , *in prep.*).

```{r session information}
rm(list=ls())
sessionInfo()
```

All codes used in this study can furthermore be found here: https://github.com/mrelnoob/ppl.tits
\
\

## 1.2. What is this document about?  

```{r data import, include=FALSE}
library(ppl.tits) # To be able to use pipes.
stab_r2plot <- targets::tar_read(RF_r2plot)
stab_varimplot <- targets::tar_read(RF_importanceplot)
# tits <- targets::tar_read(final_tdata) # I don't need data for now!
# As you can see, I import my data using 'targets'! If you are reading this report and don't know what it is or how I created and stored them, you should read the README or the _devhistory.R files available at https://github.com/mrelnoob/ppl.tits. 

.pardefault <- par() # To save the default graphical parameters (in case I want to restore them).
```
\

In the context of the *tits reproduction study* of the PubPrivLands project, we undertook a complex data processing methodology that can be divided into **four stages**:
1. **Raw data acquisition and preparation** - This initial stage consists of all *raw* data acquisition through fieldwork, GIS or remote-sensing-based data generation, and database extractions as well as their initial *cleaning* and *formatting* to produce **raw datasets** (e.g. tits reproduction observations, environmental variables). This stage is globally described in the *Materials & Methods* section of the manuscript, and is thoroughly detailed in the **ppl_log.pdf** document (French only).
2. **Clean dataset generation** - This stage includes all *cleaning*, *wrangling*, *computing* and *formatting* required to obtain a **clean exploitable dataset**. This stage is openly documented in the !!!!!

This document presents the relevant results and features of all *intermediate analyses* performed in the context of the *tits reproduction study* of the PubPrivLands project. We refer to *intermediate analyses* for all data processing meant to use ...
